---
title: "simpleRegress"
author: "Galilea Hernandez"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## Model WEIGHT ~ AGE

```{r}
agemodel <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(agemodel)
summary.lm(agemodel)
```
Best to worst:

1.NECK  ARsqrd: 0.87    RST: 43.91  
2.AGE   ARsqrd: 0.55    RST: 81.47     

The agemodel graph shows its points to be far away from the line, while the simp graph has plots closer to it.


## Model WEIGHT ~ HEADLEN

```{r}
headlenmodel <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(headlenmodel)
summary.lm(headlenmodel)
```
Best to worst:

1.NECK       ARsqrd: 0.87    RST: 43.91  
2.HEADLEN    ARsqrd: 0.69    RST: 67.81    
3.AGE        ARsqrd: 0.55    RST: 81.47    

Although the headlenmodel graph has scattered out plots it is better than the
graph of agemodel.Comparing the Adjusted R-squared being higher at 0.69 and the agemodel Adjusted R-squared is 0.55, shows it being better than it but not
high enough for simp's Adjusted R-squared of 0.87.

## Model WEIGHT ~ HEADWTH

```{r}
headwthmodel <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(headwthmodel)
summary.lm(headwthmodel)
```

Best to worst:

1.NECK       ARsqrd: 0.87    RST: 43.91  
2.HEADLEN    ARsqrd: 0.69    RST: 67.81     
3.HEADWTH    ARsqrd: 0.60    RST: 76.41   
4.AGE        ARsqrd: 0.55    RST: 81.47   


## Model WEIGHT ~ LENGTH

```{r}
lengthmodel <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(lengthmodel)
summary.lm(lengthmodel)
```

Best to worst:

1.NECK       ARsqrd: 0.87    RST: 43.91    
2.LENGTH     ARsqrd: 0.74    RST: 61.83   
3.HEADLEN    ARsqrd: 0.69    RST: 67.81      
4.HEADWTH    ARsqrd: 0.60    RST: 76.41      
5.AGE        ARsqrd: 0.55    RST: 81.47    




## Model WEIGHT ~ CHEST

```{r}
chestmodel <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(chestmodel)
summary.lm(chestmodel)
```

Best to worst:

1.CHEST.     ARsqrd: 0.92    RST: 33.08    
2.NECK       ARsqrd: 0.87    RST: 43.91    
3.LENGTH     ARsqrd: 0.74    RST: 61.83   
4.HEADLEN    ARsqrd: 0.69    RST: 67.81      
5.HEADWTH    ARsqrd: 0.60    RST: 76.41     
6.AGE        ARsqrd: 0.55    RST: 81.47     












